---
title: "p8105_hw2_ry2417"
author: "Ruiqi Yan"
date: "9/30/2021"
output: github_document
---

Load all packages needed
```{r, message = FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1

import and clean data in `Mr.Trash Wheel` sheet from `Trash-Wheel-Collection-Totals-7-2021-sept.xlsx`

```{r}
trashwheel_df <- 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", 
    range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = round(sports_balls))
```

import and clean data in `2018 Precipitation` and `2019 Precipitation` from `Trash-Wheel-Collection-Totals-7-2021-sept.xlsx`

```{r}
precipitation_2018_df <-
  read_excel(
    "data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14"
  ) %>% 
  janitor::clean_names() %>% 
  rename(precipitation_inch = total) %>% 
  filter(!is.na(precipitation_inch)) %>% 
  mutate(year = "2018")

precipitation_2019_df <-
  read_excel(
    "data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14"
  ) %>% 
  janitor::clean_names() %>% 
  rename(precipitation_inch = total) %>% 
  filter(!is.na(precipitation_inch)) %>% 
  mutate(year = "2019")
```

combine two precipitation data frames and convert month to a character variable 

```{r}
precipitation_all_df <-
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year)
```


describe these two data sets

```{r summary_tables_trashwheel}
some_example_trashwheel <- 
  trashwheel_df %>% 
  select(dumpster, date, weight_tons, volume_cubic_yards) %>% 
  head()

year_summary <- 
  trashwheel_df %>% 
  group_by(year) %>% 
  summarise(
    total_weight_tons = sum(weight_tons), 
    average_weight_tons = mean(weight_tons), 
    total_volume_cubic_yards = sum(volume_cubic_yards),
    average_volume_cubic_yards = mean(volume_cubic_yards)
  )

median_sport_balls_2019 <- 
  trashwheel_df %>% 
  filter(year == 2019) %>% 
  pull(sports_balls) %>% 
  median()

some_example_precipitation <- 
  precipitation_all_df %>% 
  head()

max_precipitation <- 
  precipitation_all_df %>% 
  filter(precipitation_inch == max(precipitation_inch)) %>% 
  select(month, year) %>% 
  unlist()

min_precipitation <- 
  precipitation_all_df %>% 
  filter(precipitation_inch == min(precipitation_inch)) %>% 
  select(month, year) %>% 
  unlist()

sum_precipitation_2018 <- 
  precipitation_all_df %>% 
  filter(year == "2018") %>% 
  pull(precipitation_inch) %>% 
  sum()

sum_precipitation_2019 <- 
  precipitation_all_df %>% 
  filter(year == "2019") %>% 
  pull(precipitation_inch) %>% 
  sum()
```

There are `r nrow(trashwheel_df)` observations and `r ncol(trashwheel_df)` 
variables in `Mr.Trash Wheel`. The variables include: ``r colnames(trashwheel_df)``. 
Some examples of key variables as 
`r knitr::kable(some_example_trashwheel)` The year of the data is ranging from 
`r trashwheel_df %>% pull(year) %>% min()` to 
`r trashwheel_df %>% pull(year) %>% max()`. The total number of cigarette butts 
are $`r trashwheel_df %>% pull(cigarette_butts) %>% sum()`$; the total of glass 
bottles are $`r trashwheel_df %>% pull(glass_bottles) %>% sum()`$; the total of 
plastic bottles are $`r trashwheel_df %>% pull(plastic_bottles) %>% sum()`$. 
The total and average weight and volume of each year are 
`r  knitr::kable(year_summary)`
The median number of sports balls in the dumpster in 2019 is 
`r median_sport_balls_2019`. 2018 has most trash and 2019 has least trash.\
There are `r nrow(precipitation_all_df)` observations and `r ncol(precipitation_all_df)` 
variables in the combination of `2018 precipitation` and `2019 precipitation`. 
The variables include: ``r colnames(precipitation_all_df)``. Some examples of key
variables as `r knitr::kable(some_example_precipitation)` 
The month with most precipitation is `r max_precipitation` and the month with 
least precipitation is `r min_precipitation`. The total precipitation in 2018 is
`r sum_precipitation_2018` inches. The total precipitation in 2019 is 
`r sum_precipitation_2019` inches.

## Problem 2 

import and clean data in `pols-month.csv`

```{r, message = FALSE}
pols_df <- 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    mon, 
    into = c("year", "month", "day"), 
    sep = "-", 
    convert = TRUE
  ) %>% 
  mutate(
    month = month.name[month],
    president = ifelse(prez_dem == 1, "dem", "gop")
  ) %>% 
  select(-prez_dem, -prez_gop, -day)
  
```

import and clean data in `snp.csv`

```{r, message = FALSE}
snp_df <- 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    date, 
    into = c("month", "day", "year"), 
    sep = "/", 
    convert = TRUE
  ) %>% 
  mutate(
    year = ifelse(year <= 15, year + 2000, year + 1900),
    month = month.name[month],
  ) %>% 
  select(-day) %>% 
  relocate(year, month) %>% 
  arrange(year, month)
```

import and clean data in `unemployment.csv`

```{r, message = FALSE}
unemployment_df <-
  read_csv("data/unemployment.csv") %>% 
  rename_with(~ month.name, month.abb) %>% 
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "percent_unemployment"
  ) %>% 
  janitor::clean_names()
```

join the data sets by merging `snp` into `pols`, and join the `unemployment` to 
the result data frame.

```{r, message = FALSE}
merging_df <-
  pols_df %>% 
  left_join(snp_df, by = c("year", "month")) %>% 
  left_join(unemployment_df, by = c("year", "month"))
```

describe these data sets
```{r}
max_stock_month <-
  merging_df %>% 
  filter(close == max(close, na.rm = TRUE)) %>% 
  select(month, year, president) %>% 
  unlist()

min_stock_month <-
  merging_df %>% 
  filter(close == min(close, na.rm = TRUE)) %>% 
  select(month, year, president) %>% 
  unlist()

max_unemployment_month <-
  merging_df %>% 
  filter(percent_unemployment == max(percent_unemployment, na.rm = TRUE)) %>% 
  select(month, year, president) %>% 
  unlist()

min_unemployment_month <-
  merging_df %>% 
  filter(percent_unemployment == min(percent_unemployment, na.rm = TRUE)) %>% 
  select(month, year, president) %>% 
  unlist()
  
some_example_merging <- 
  merging_df %>% 
  select(year, month, president, close, percent_unemployment) %>% 
  filter(!is.na(close)) %>% 
  head()
```

Data in `pols` has `r nrow(pols_df)` observations of `r ncol(pols_df)` variables.
The year is ranging from `r pols_df %>% pull(year) %>% min()` to 
`r pols_df %>% pull(year) %>% max()`.\
Data in `snp` has `r nrow(snp_df)` observations of `r ncol(snp_df)` variables.
The year is ranging from `r snp_df %>% pull(year) %>% min()` to 
`r snp_df %>% pull(year) %>% max()`.\
Data in `unemployment` has `r nrow(unemployment_df)` observations of 
`r ncol(unemployment_df)` variables.
The year is ranging from `r unemployment_df %>% pull(year) %>% min()` to 
`r unemployment_df %>% pull(year) %>% max()`.\
The result data frame has `r nrow(merging_df)` observations of 
`r ncol(merging_df)` variables. The year is ranging from 
`r snp_df %>% pull(year) %>% min()` to 
`r snp_df %>% pull(year) %>% max()`.\
The variables of merging data frame include ``r names(merging_df)``.
Some examples of key variables are `r knitr::kable(some_example_merging)` The 
months with highest unemployment percentage are `r max_unemployment_month[c(1,3)]` 
and `r max_unemployment_month[c(2,4)]`, when the president was republican. The 
months with lowest unemployment percentage is
`r min_unemployment_month[c(1,3)]` and `r min_unemployment_month[c(2,4)]`, 
when the president was republican. The month with 
the highest closing values of the S&P stock  is `r max_stock_month[-3]` 
when the president was democratic. The 
month with the lowest closing values of the S&P stock  is `r min_stock_month[-3]` 
when the president was republican.

## Problem 3

import and clean data in `Popular_Baby_Names`.

```{r, message = FALSE}
babynames_df <- 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),  
    ethnicity = str_to_title(ethnicity),
    childs_first_name = str_to_title(childs_first_name)
  ) %>% 
  mutate(
    ethnicity = 
      recode(
      ethnicity,
      "Black Non Hisp" = "Black Non Hispanic", 
      "White Non Hisp" = "White Non Hispanic",
      "Asian And Paci" = "Asian And Pacific Islander"
    )
  ) %>% 
  distinct()
```

The table showing the rank in popularity of the name “Olivia” as a female baby 
name over time respect to different ethnicity

```{r, message = FALSE}
babynames_df %>% 
  filter(childs_first_name == "Olivia", gender == "female") %>% 
  select(rank, year_of_birth, ethnicity) %>%
  arrange(year_of_birth) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>% 
  knitr::kable()
```

The table showing the most popular name among male children over time respect to
different ethnicity

```{r}
babynames_df %>% 
  filter(gender == "male", rank == 1) %>% 
  select(childs_first_name, year_of_birth, ethnicity) %>% 
  arrange(year_of_birth) %>% 
  pivot_wider(
    names_from = (year_of_birth),
    values_from = childs_first_name
  ) %>% 
  knitr::kable()
```

Create a scatter plot showing the number of children with a name (y axis) 
against the rank in popularity of that name (x axis) for white non-hispanic male 
children born in 2016 

```{r male_name_2016_scatterplot, fig.width = 6, fig.height = 6}
babynames_df %>% 
  filter(
    gender == "male", 
    year_of_birth == 2016,
    ethnicity == "White Non Hispanic"
  ) %>% 
  select(count, rank) %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() +
  labs(
    x = "rank of the name in 2016", 
    y = "the number of children with the name in 2016"
  )
```