---
title: "p8105_hw2_ry2417"
author: "Ruiqi Yan"
date: "9/30/2021"
output: github_document
---

Load all packages needed
```{r, message = FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1

import and clean data in `Mr.Trash Wheel` sheet from `Trash-Wheel-Collection-Totals-8-6-19.xlsx`

```{r}
trashwheel_df <- 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = round(sports_balls))
```

import and clean data in `2018 Precipitation` and `2019 Precipitation` from `Trash-Wheel-Collection-Totals-8-6-19.xlsx`

```{r}
precipitation_2018_df <-
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14"
  ) %>% 
  janitor::clean_names() %>% 
  rename(precipitation_inch = total) %>% 
  filter(!is.na(precipitation_inch)) %>% 
  mutate(year = "2018")

precipitation_2019_df <-
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14"
  ) %>% 
  janitor::clean_names() %>% 
  rename(precipitation_inch = total) %>% 
  filter(!is.na(precipitation_inch)) %>% 
  mutate(year = "2019")
```

combine two precipitation data frames and convert month to a character variable 

```{r}
precipitation_all_df <-
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year)
```


describe these two data sets

```{r summary_tables_trashwheel}
some_example_trashwheel <- 
  trashwheel_df %>% 
  select(dumpster, date, weight_tons, volume_cubic_yards) %>% 
  head()

year_summary <- 
  trashwheel_df %>% 
  group_by(year) %>% 
  summarise(
    total_weight_tons = sum(weight_tons), 
    average_weight_tons = mean(weight_tons), 
    total_volume_cubic_yards = sum(volume_cubic_yards),
    average_volume_cubic_yards = mean(volume_cubic_yards)
  )

median_sport_balls_2017 <- 
  trashwheel_df %>% 
  filter(year == 2017) %>% 
  pull(sports_balls) %>% 
  median()

some_example_precipitation <- 
  precipitation_all_df %>% 
  head()

max_precipitation <- 
  precipitation_all_df %>% 
  filter(precipitation_inch == max(precipitation_inch)) %>% 
  select(month, year) %>% 
  unlist()

min_precipitation <- 
  precipitation_all_df %>% 
  filter(precipitation_inch == min(precipitation_inch)) %>% 
  select(month, year) %>% 
  unlist()

sum_precipitation_2018 <- 
  precipitation_all_df %>% 
  filter(year == "2018") %>% 
  pull(precipitation_inch) %>% 
  sum()

sum_precipitation_2019 <- 
  precipitation_all_df %>% 
  filter(year == "2019") %>% 
  pull(precipitation_inch) %>% 
  sum()
```

There are `r nrow(trashwheel_df)` observations and `r ncol(trashwheel_df)` 
variables in `Mr.Trash Wheel`. The variables include: ``r colnames(trashwheel_df)``. 
Some examples of key variables as 
`r knitr::kable(some_example_trashwheel)`. The year of the data is ranging from 
`r trashwheel_df %>% pull(year) %>% min()` to 
`r trashwheel_df %>% pull(year) %>% max()`. The total number of cigarette butts 
are $`r trashwheel_df %>% pull(cigarette_butts) %>% sum()`$; the total of glass 
bottles are $`r trashwheel_df %>% pull(glass_bottles) %>% sum()`$; the total of 
plastic bottles are $`r trashwheel_df %>% pull(plastic_bottles) %>% sum()`$. 
The total and average weight and volume of each year are 
`r  knitr::kable(year_summary)`. 
The median number of sports balls in a dumpster in 2017 is 
`r median_sport_balls_2017`. 2018 has most trash and 2019 has least trash.\
There are `r nrow(precipitation_all_df)` observations and `r ncol(precipitation_all_df)` 
variables in the combination of `2018 precipitation` and `2019 precipitation`. 
The variables include: ``r colnames(precipitation_all_df)``.Some examples of key
variables as `r knitr::kable(some_example_precipitation)` 
The month with most precipitation is `r max_precipitation` and the month with 
least precipitation is `r min_precipitation`. The total precipitation in 2018 is
`r sum_precipitation_2018` inches. The total precipitation in 2019 is 
`r sum_precipitation_2019` inches.

## Problem 2 

import and clean data in `pols-month.csv`

```{r, message = FALSE}
pols_df <- 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    mon, 
    into = c("year", "month", "day"), 
    sep = "-", 
    convert = TRUE
  ) %>% 
  mutate(
    month = month.name[month],
    president = ifelse(prez_dem == 1, "dem", "gop")
  ) %>% 
  select(-prez_dem, -prez_gop, -day)
  
```

import and clean data in `snp.csv`

```{r, message = FALSE}
snp_df <- 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    date, 
    into = c("month", "day", "year"), 
    sep = "/", 
    convert = TRUE
  ) %>% 
  mutate(
    year = ifelse(year <= 15, year + 2000, year + 1900),
    month = month.name[month],
  ) %>% 
  select(-day) %>% 
  relocate(year, month) %>% 
  arrange(year, month)
```